---
title: "Fastlink"
author: "Haocheng Qin, Kewei Xu, Zaolin Zhang, Chuhan Guo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Methodology

Literature: https://imai.fas.harvard.edu/research/files/linkage.pdf

To conclude the methodology section of the paper, the authors developed a fast and scalable probabilistic model designed to merge large-scale administrative records more effectively than traditional deterministic methods. This model, called fastLink, addresses issues such as missing data, measurement error, and the uncertainty inherent in merging processes, which are common in social science research. By incorporating auxiliary information, such as name frequency or migration rates, and allowing for robust simulation studies, the proposed methodology significantly outperforms deterministic approaches. Furthermore, it offers an open-source solution to merge data sets efficiently while providing tools for post-merge analyses that account for the uncertainty of the matching process​

## Analysis

Uncertainty Analysis

Scalability

Empirical Validation

## Package Realization

Repository: https://github.com/kosukeimai/fastLink

Example: https://imai.fas.harvard.edu/research/files/turnout.pdf

## Implementation 

Dataset: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/2NNA4L

```{r}
suppressMessages(require("fastLink"))
suppressMessages(require("plyr"))
data <- read.delim("cces2016voterval.tab")
summary(data)
```

