---
title: "Demo"
author: "Haocheng, Kewei, Zaolin, Chuhan"
output:
  powerpoint_presentation:
    reference_doc: "reference.pptx" 
  beamer_presentation:
    theme: AnnArbor
    colortheme: seahorse
    fonttheme: structurebold
  slidy_presentation: default
  ioslides_presentation: default
bibliography: references.bib
link-citations: yes
csl: ieee.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

+ FastLink is a scalable entity resolution methodology designed for merging large-scale datasets.
+ Advantages
  - Addresses challenges such as missing data, measurement errors, and uncertainty in the merging process
  - Provide more flexibility by using auxiliary information (e.g., name frequency, migration rates)
  - Utilize a probabilistic match score for more accurate linking, even with incomplete or imprecise data
  - Scalable and capable of handling millions of records, making it efficient in terms of speed and accuracy
+ Limitations
  - Less effective with long strings (e.g., full names, long addresses) due to variations and typographical errors without advanced string-matching algorithms

## Methodology

- General settings
$$
\begin{aligned}
&\text{Latent Matches } && M_{i j} \stackrel{\text { i.i.d. }}{\sim} \operatorname{Bernoulli}(\lambda)\\
&\text{Distance } && \gamma_k(i, j) \mid M_{i j}= m \stackrel{\text { indep. }}{\sim} \operatorname{Discrete}\left(\pi_{k m}\right) \\
&\text{Missing Indicator } &&\delta_k(i, j) \perp \gamma_k(i, j) \mid M_{i j}\\
&\text{Probability }&&\xi_{i j} :=\operatorname{Pr}\left(M_{i j}=1 \mid \delta(i, j), \gamma(i, j)\right)
\end{aligned}
$$
+ Quite similar to Fellegi-Sunter[@fellegi1969]
+ Capable pf dealing MAP missings

## Methodology

+ Probabilistic model, see getPosterior.R
$$
\begin{aligned}
\xi_{i j}  =\frac{\lambda \prod_{k=1}^K\left(\prod_{\ell=0}^{L_k-1} \pi_{k \ell \ell}^{1\left\{\gamma_k(i, j)=\ell\right\}}\right)^{1-\delta_k(i, j)}}{\sum_{m=0}^1 \lambda^m(1-\lambda)^{1-m} \prod_{k=1}^K\left(\prod_{\ell=0}^{L_k-1} \pi_{k m \ell}^{1\left\{\gamma_{\ell}(i, j)=\ell\right\}}\right)^{1-\delta_k(i, j)}}
\end{aligned}
$$
+ Nice for evaluation and post-merge analysis
$$
X^*_i=\sum_{j=1}^{N_{\mathcal{B}}} \xi_{i j} X_j / \sum_{j=1}^{N_{\mathcal{B}}} \xi_{i j}
$$

## Methodology

+ Nice for calculation
  - Likelihood and boosting with EM steps[@winkler2000]
$$
\begin{aligned}
\lambda & =\frac{1}{N_{A} N_{B}} \sum_{i=1}^{N_{A}} \sum_{j=1}^{N_{B}} \xi_{i j} \\
\pi_{k m \ell} & =\frac{\left.\sum_{i=1}^{N_{A}} \sum_{j=1}^{N_{B}} \mathbf{1}\left\{\gamma_k(i, j)=l\right)\right\}\left(1-\delta_k(i, j)\right) \xi_{i j}^m\left(1-\xi_{i j}\right)^{1-m}}{\sum_{i=1}^{N_{A}} \sum_{j=1}^{N_{B}}\left(1-\delta_k(i, j)\right) \xi_{i j}^m\left(1-\xi_{i j}\right)^{1-m}}
\end{aligned}
$$
  - logemlink.R and emlinkMARmov.R

## Package Implement

The required package for FastLink is called “fastLink”:

```{r, eval=FALSE}
install.packages("fastLink")
```

Install the most recent version of "fastLink" package (version 0.6):

```{r, eval=FALSE}
library(devtools)
install_github("kosukeimai/fastLink",dependencies=TRUE)
```

```{r, eval=FALSE}
## Load the package and data
library(fastLink)
```

Tutorial Link: https://github.com/kosukeimai/fastLink

## Package Implement

```{r, eval=FALSE}
matches.out <- fastLink(
  dfA = dfA, dfB = dfB, 
  varnames = c("given_name", "surname", "address_1", "suburb"),
  stringdist.match = c("given_name", "surname"),
  partial.match = c("given_name", "surname"),
  return.all = TRUE
)
```

The merged dataset can be accessed using the getMatches() function:

```{r, eval=FALSE}
matched_dfs <- getMatches(
  dfA = dfA, dfB = dfB, 
  fl.out = matches.out, threshold.match = 0.85
)
```

## Package Implement

- Preprocessing Matches via Blocking: The blockData() function can block two datasets using one or more variables and various blocking techniques.

- Using Auxiliary Information to Inform fastLink: The algorithm could also incorporate auxiliary information on migration behavior to inform the matching of datasets over time.

- Aggregating Multiple Matches Together: The algorithm can also aggregate multiple matches into a single summary using the aggregateEM() function.

- Random Sampling with fastLink: The algorithm allows us to run the matching algorithm on a randomly selected smaller subset of data to be matched and then apply those estimates to the full sample of data.

- Finding Duplicates within a Dataset via fastLink: The algorithm uses the probabilistic match algorithm to identify duplicated entries.

## Dataset

+ Two sets of datasets explored
  - Products on Amazon & Google
  - fictious dataset from Freely Extensible Biomedical Record Linkage
+ Empirical evidence of reduced effectiveness with long strings (e.g., product descriptions)
  - only 75 matches for Amazon (>1000 rows) & Google (>3000 rows) product datasets
+ Avoid overly broad matching criteria / using too many variables
  - Inflate match rate (even over 100%), underestimate False Discovery Rate (FDR) and False Non-match Rate (FNR)

("https://github.com/ShimizuYukii/Entity-Resolution/blob/main/RS%20data/over_match.png")

## Dataset

+ choose more informative variable could reduce the use of matching variables
  - name + soc_sec_id vs name + address + suburb
+ number of matches & jaccard index vs threshold probabilities
  - jaccard index = intersection size / union size

("https://github.com/ShimizuYukii/Entity-Resolution/blob/main/RS%20data/address1%2Bsuburb.png")
("https://github.com/ShimizuYukii/Entity-Resolution/blob/main/RS%20data/soc_sec_id.png")
("https://github.com/ShimizuYukii/Entity-Resolution/blob/main/RS%20data/num_match%20vs%20prob.png")
("https://github.com/ShimizuYukii/Entity-Resolution/blob/main/RS%20data/jaccard%20vs%20prob.png")



## References

<div id="refs"></div>
